---
title: "Summary"
output:
  pdf_document: default
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_dir=file.path(getwd(), 'data/reports')) })
---

```{r config, include=FALSE}
knitr::opts_knit$set(root.dir = file.path(getwd(), '..'))
options(digits = 4)
knitr::opts_chunk$set(comment = NA)
```

```{r setup, include=FALSE}
library(tidyverse)
library(ggbeeswarm)
library(scales)
library(DBI)

source("src/stats_utils/utils.R")
```

```{R load_enrolments, echo=FALSE}
grades <- (
    fetch("SELECT username, course, semester, grade, sat_exam from Grade")
    |> mutate(
         course = relabel_courses(course),
         sat_exam = relabel_exam(sat_exam),
         semester = relabel_semester(semester),
         grade = grade * 100
       )
)
```

## Summary

The study was divided over two semesters and two courses. The enrolments for
each course and semester are shown below.

```{R plot_enrolments, echo=FALSE, fig.height=3, fig.width=5.7}
(
    ggplot(grades, aes(y = fct_rev(course), fill = fct_rev(sat_exam)))
    + geom_bar(color = "black")
    + labs(
        title = "Students enrolled by Semester",
        x = "Number of Students (Stacked)", y = "",
        fill = "Sat Exam?"
      )
    + facet_wrap(vars(semester), ncol = 1, scales = "free")
    + custom_theme()
    + scale_fill_grey(start = 0.8, end = 0.4)
    + scale_x_continuous(breaks = breaks_pretty())
)
```

```{r tabulate_enrolments, echo=FALSE}
t(
  grades
  |> group_by(course, semester)
  |> summarise(
       n = n_distinct(username),
       `sat exam` = sum(sat_exam == "Sat Exam"),
       `sat exam%` = `sat exam` / n * 100,
       `did not sit exam` = sum(sat_exam == "Did Not Sit"),
       `did not sit exam%` = `did not sit exam` / n * 100,
       .groups = "drop"
     )
)
```

It is common for students to withdraw/"drop out" before the end of a course;
however, data for who has withdrawn is not available. As a proxy for "still
enrolled students", this study considers only those students who sat the final
exam for each course. This will exclude some students who could not sit the
final exams for legitimate reasons, but from the above tables, this excludes
only about 6% of the data.

```{r filter_enrolments}
grades <- grades |> filter(sat_exam == "Sat Exam")
```

The overall grades for students who completed each course are below. This is the
unweighted average for students test and exam scores—ignoring missing test
scores.

```{R plot_grades, echo=FALSE}
summary <- (
  grades
  |> group_by(course , semester)
  |> summarise(
       mean = mean(grade),
       median = median(grade),
       n = n(),
       .groups = "drop"
     )
)

(
  ggplot(grades, aes(x = grade))
  + geom_density(fill = "grey90", alpha = 0.8, linewidth = 0.7)
  + labs(
      title = "Density of Student Grades by Course and Semester",
      x = "Grade (/100)", y = "Density"
    )
  + geom_vline(
      aes(xintercept = mean, linetype = "mean"),
      data = summary, 
      key_glyph = "path"
    )
  + geom_vline(
      aes(xintercept = median, linetype = "median"),
      data = summary,
      key_glyph = "path"
    )
  + geom_text(
      aes(label = paste0("n=", n), x = -Inf, y = Inf),
      data = summary,
      hjust = -0.2, vjust = 2,
      fontface = "bold"
    )
  + facet_grid(semester ~ course)
  + custom_theme(panel.spacing = unit(1, "lines"))
  + scale_x_continuous(limits = c(0, 100), expand = c(0, 0))
  + scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
  + scale_linetype_manual(
      name = "Summary", 
      values = c(mean = "solid", median = "dashed")
    )
)
```

```{R tabulate_grades, echo=FALSE}
t(
  grades
  |> group_by(course, semester)
  |> summarise(
       n = n_distinct(username),
       `mean grade` = mean(grade),
       `median grade` = median(grade),
       `stddev` = sd(grade),
       .groups = "drop"
     )
)
```

```{R load_cardsorts, echo=FALSE}
data <- (
  fetch("
      SELECT Attempt.question,
             Attempt.username,
             Attempt.course,
             Attempt.semester,
             Attempt.idx,
             Attempt.grade as mark,
             Attempt.is_admissible,
             Attempt.is_genuine,
             Grade.grade as grade
      FROM Attempt
               JOIN Cardsort ON Attempt.question = Cardsort.name
               LEFT JOIN Grade USING (username, course, semester)
      WHERE Grade.sat_exam = 1
  ")
  |> mutate(
         course = relabel_courses(course),
         semester = relabel_semester(semester),
         question = relabel_questions(question),
         grade = grade * 100,
         mark = mark * 100
       )
)

num_attempts <- (
  data
  |> group_by(question, username, course, semester)
  |> summarise(
       attempts = max(idx) + 1,
       is_genuine = any(idx == 0 & is_genuine == 1),
       .groups = "drop"
     )
  |> group_by(question, course, semester)
  |> mutate(is_outlier = is_outlier(attempts))
  |> left_join(grades, by = join_by(course, semester, username))
)
```

```{R echo=FALSE, fig.width=7.5, message=FALSE}
(
  ggplot(num_attempts, aes(x = question, y = attempts))
  + geom_boxplot(notch = TRUE)
  + geom_quasirandom(width = 0.15, size = 0.3)
  + labs(
      title = "Raw Number of Attempts Per Question",
      x = "Question", y = "Number of Attempts"
    )
  + facet_grid(semester ~ course, scales = "free")
  + custom_theme(axis.text = element_text(size = 8))
  + scale_x_discrete(labels = label_wrap(width = 15))
)
```

A few observations after looking at the data:

-   Outliers: Some students had an unusually large number of attempts. It is
    assumed these students are attempting to brute-force solutions
-   Strategy: Some students did not move all cards before clicking "check". This
    might indicate a form of brute-forcing where students continually check
    questions before constructing a full solution.

Below, this data is tabulated.
    
```{r echo=FALSE}
t(
  num_attempts
  |> summarise(
       n = n(),
       `genuine attempt` = sum(!is_outlier & is_genuine),
       `genuine attempt%` = `genuine attempt` / n * 100,
       `not genuine` = sum(is_outlier | !is_genuine),
       `not genuine%` = `not genuine` / n * 100,
       `outlier only` = sum(is_outlier & is_genuine),
       `didn't move all only` = sum(!is_outlier & !is_genuine),
       `Both an outler and didn't move all` = sum(is_outlier & !is_genuine),
       .groups = "drop"
     )
  |> arrange(course, semester, question)
)
```

It is assumed that students who do not give genuine attempts to questions do so as they are weak students.

```{R echo=FALSE, fig.width=9, message=FALSE}
(
  ggplot(num_attempts, aes(x = question, y = grade, fill = is_outlier))
  + geom_boxplot(notch = TRUE)
  + geom_quasirandom(width = 0.12, size = 0.2, dodge.width = 0.75)
  + labs(
      title = "Grades By Question and Outlier Status",
      x = "Question", y = "Grade (/100)",
      fill = "Is Outlier?"
    )
  + facet_grid(semester ~ course, scales = "free")
  + custom_theme(axis.text = element_text(size = 8))
  + scale_x_discrete(labels = label_wrap(width = 15))
  + scale_fill_grey(start = 1, end = 0.6)
)
```

Outliers generally seem to be the weakest students. For further analysis, these
students are removed as it is assumed that to achieve a number of attempts so
extreme they are not genuinely attempting the problm.

```{R}
num_attempts <- num_attempts |> filter(!is_outlier)
```

Next, a plot showing the numbers of attempts for students who did/did not move
all cards in their first attempt is shown.

```{R echo=FALSE, fig.width=9, message=FALSE}
(
  ggplot(num_attempts, aes(x = question, y = grade, fill = is_genuine))
  + geom_boxplot(notch = TRUE)
  + geom_quasirandom(width = 0.12, size = 0.2, dodge.width = 0.75)
  + labs(
      title = "Grade By Question and Strategy",
      x = "Question", y = "Grade",
      fill = "Moved all Cards?"
    )
  + facet_grid(semester ~ course, scales = "free")
  + custom_theme(axis.text = element_text(size = 8))
  + scale_x_discrete(labels = label_wrap(width = 15))
  + scale_y_continuous(limits = c(0, 100))
  + scale_fill_grey(start = 1, end = 0.6)
)
```

Students not giving genuine attempts also seem to be relatively weak. They are removed.

```{r}
num_attempts <- num_attempts |> filter(is_genuine)
```

Now, the number of attempts per question looks like:

```{R echo=FALSE, fig.width=7.5, message=FALSE}
(
  ggplot(num_attempts, aes(x = question, y = attempts))
  + geom_boxplot(notch = TRUE)
  + geom_quasirandom(width = 0.15, size = 0.3)
  + labs(
      title = "Number of Attempts Per Question for Genuine Attempts",
      x = "Question", y = "Number of Attempts"
    )
  + facet_grid(semester ~ course, scales = "free")
  + custom_theme(axis.text = element_text(size = 8))
  + scale_x_discrete(labels = label_wrap(width = 15))
)
```
